A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-28 manuscript.Rmd
fix botched merge

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
source(here::here("R", "simulation_funs.R"))
```

@@

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research hinges on its objectivity. Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper]. Therefore, reproducibility has long been considered integral to empirical research.
<!---
You tend to use quite long sentences with many subclauses. While this has a long tradition in the german language, it is considered bad style in english. Also it often leads to minor mistakes like dangling modifiers.
-->
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.
First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
This workflow builds on several technical solutions originating in the software engineering community. 
We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
Second, we discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
The resulting dynamic document closely resembles the final manuscript, but uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
Once the empirical data are analyzed, the Results are updated, and the Discussion is written.

@@

Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
<!---
I am not shure about the word artifact in general. Is it common terminology?
-->
However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If research artifacts are available but not reproducible, their added value is limited.

This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is more likely to be adopted.
Our approach is to promote broad adoption of easy-to-use practices (a "low threshold"), and ensure that these practices are compatible with more complex rigorous solutions (a "high ceiling").
As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.

@@

1.	**Multiple inconsistent versions of code, data, or both**; for example
  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
<!-- CJ 5. Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
<!--AB Caspar's 5th point is important. In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken. Can we merge this into one new point #3? I will commit a suggested change of point #3 here (see above);.-->

We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses. It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph. The opening paragraph should at least:
1) Introduce the topic area
2) Introduce the problem/knowledge gap
3) Introduce how you will fill this knowledge gap-->

@@

Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
Git is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit "knows" its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
Git has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
Collaborating via Git requires the repository to be uploaded somewhere.
We recommend GitHub to host Git repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.

Second, we rely on **dynamic document generation**.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
Typically, results are manually copied and pasted from one software another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.

One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

To automatically resolve dependencies between project files, we rely on **dependency tracking**.
In essence, researchers provide a collection of computational instructions (_recipes_) that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).

We recommend using **Make** for dependency tracking because it is language independent.
The following hypothetical example illustrates the utility of Make.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
A Makefile for this project could look like this:

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

**Docker** is an example of containerization.
It provides all software used in an analysis, including the operating system itself.
(The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.)
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend storing a complete image of the software environment besides the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)); dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).    
While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

@@

The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:

```{r, eval=FALSE}
# repro is not on CRAN yet
install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
```

An example of the assistance that `repro` provides are the check functions.
They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :

@@

They detect the users' system state and explain if further action is needed.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:

@@

The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
If necessary, follow the instructions of repro to install them until the check functions are all fine.

The first concept we want to introduce is version control with Git.
Git tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

@@

You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```

Now the file `private.md` will not be committed to Git and will not be made public.

After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:

@@

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
The easiest way to do this is to use the RStudio Git pane.
Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and share the files, you can publicly upload those changes to a public repository.
To synchronize your local repository with a remote repository from the GitHub service, issue the following command:

@@

@@

If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:

Now that you have created a project that is under version control, we continue with adding dynamic document generation.
Adynamic document has three elements:

1. Text in a natural language
2. Executable code 
3. Metadata

The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

First, preregistrations are plagued by non-reproducibility like any other research outcome.
@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).

>>>>

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
```

@@

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper]
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.
First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
This workflow builds on several technical solutions originating in the software engineering community. 
We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics. 
After the empirical data are analyzed, the Results are updated, and the Discussion is written.

@@

Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
If research artifacts are available but not reproducible, their added value is limited.

This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is more likely to be adopted.
Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.

@@

1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
<!-- CJ 5. Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
<!--AB Caspar's 5th point is important. In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken. Can we merge this into one new point #3? I will commit a suggested change of point #3 here (see above);.-->

We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph. The opening paragraph should at least:
1) Introduce the topic area
2) Introduce the problem/knowledge gap
3) Introduce how you will fill this knowledge gap-->

@@

Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
Git is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit "knows" its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
Git has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
Collaborating via Git requires the repository to be uploaded somewhere.
We recommend GitHub to host Git repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.

Second, we rely on **dynamic document generation**.
The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].

Dynamic document generation integrates both steps.
Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing.
It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.

One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.

To automatically resolve dependencies between project files, we rely on **dependency tracking**.
In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).

We recommend using **Make** for dependency tracking because it is language independent.
The following hypothesical example illustrates the utility of Make.
Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
A Makefile for this project could look like this:

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

**Docker** is an example of containerization.
It provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).    
While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

@@

One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to apply to all users. Your specific situation might involve additional, more specialized steps. After completing the tutorial, you should be able to customize your workflow accordingly.
The first step is to install the required software.
Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
It will assist you while you follow the tutorial.

```{r, eval=FALSE}
# repro is not on CRAN yet
options(
  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
            CRAN = 'https://cloud.r-project.org')
)
install.packages('repro')
```

An example of the assistance that `repro` provides are the check functions.
These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :

@@

These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:

@@

The messages from `repro` try to take the user by the hand to help them solve problems.
These messages are adjusted to your specific operating system and installed dependencies.
Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
If necessary, follow any instructions presented until all checks pass.

## Implementing version control

Now that your system is set up for this tutorial, we will introduce you to version control with Git.
Git tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

@@

Git tracks any files that you manually add to the repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```
Now the file `private.md` will not be committed to Git and will not be made public.

Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
When you are ready, you can begin tracking your files using Git by running:

@@

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
One way to do this is through the visual user interface in the RStudio Git pane.
Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
This stores a snapshot of the current state of the project.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and/or share the files online, you can push your local repository to a remote repository.
While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
To upload the project publically to GitHub use:

@@

To upload it privately:

@@

## Dynamic document generation

Now that you have created a version controlled project, we will proceed with dynamic document generation.
A dynamic document has three elements:

1. Text in natural language (prose; e.g., a scientific paper or presentation)
2. Executable code (e.g., analyses)
3. Metadata (e.g. title, authors, document format)

R Markdown is a type of dynamic document well-suited to the RStudio user interface.
The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

First, preregistrations are plagued by non-reproducibility like any other research outcome.
@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
